Active Learning for SAT-Solver Benchmarking

### 2022-10-06 paper/paper.tex
Address comments in sections 1 to 4

Note that we know the runtimes of all solvers, except the new one, on all instances.
There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
%JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking? "similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
%JB: we should rename "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
%JB: for what do we measure cost here?
% (1) solving new-solver-problem once? -> apart from more runtimes to be acquired beforehand, is running our framework affected by increasing the set of existing solvers, and how does this depend on the configuration of our framework, i.e., sampling/selection/stopping? (e.g., Wilcoxon stopping criterion iterates over set of existing solvers, showing linear complexity, while fixed-subset stopping is independent from number of existing solvers)
% (2) solving new-solver-problem for each solver out of a set of solvers? -> is a rather theoretical scenario, because our algorithm itself assumes we know runtimes of all but one solver, so in practice we cannot throw in a bunch of new solvers and get a ranking over all of them
We avoid comparing pairs of solvers, which would scale quadratically.
%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if binary-search-/recursive-sorting-like procedure used
Since a new solver idea is rarely best at first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR-2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)

For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
%JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
%JB: as commented above, we should either remove this criterion or show it in the table

@@

Benchmark-set selection is an important issue for solver competitions, e.g., the SAT Competition.
In such competitions, the organizers define the rules for composing the corresponding benchmarks.
Balint et al.~\cite{balint2015overview} provide an overview of benchmark-selection criteria in different solver competitions.
%JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al. (2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
M{\i}s{\i}r~\cite{misir2021benchmark} presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances.
The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed criteria
Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.

@@

\begin{table}[tbp]
  \centering
  \caption{
  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
  }
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

\subsubsection{SAT-solver configuration.}
Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
The goal is to tune SAT-solvers for a given sub-domain of problem instances.

SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
%JB: discussion of SMAC is too long, we should cut it a bit
Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
There are mainly two similarities.
First, both approaches are model-based.
While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already known runtimes of instances and runtime information on already known solvers.
Second, both approaches employ some kind of selection strategy that is based upon the previously mentioned model.
Theirs selects configurations rather than instances though.
Furthermore, Hutter et. al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with little data points) to select the next configuration for which to acquire runtime information.
SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.

While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
The main difference is that Matricon et. al. compare a solver to another one, while we compare to a set of existing solvers.
To the best of our knowledge, we are the first to tackle the latter problem.
%JB: again, I think there are too many details for describing this single approach

Given a pair of SAT solvers, Matricon et. al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et. al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method.
Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
%JB: in our evaluation, we should state whether our results are consistent to this or not

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment in intro

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure if this dilemma arises here as well; dilemma is typical for searching one optimal solution in continuous space, where there is an incentive to exploit neighborhood in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well

>>>>

Note that we know the runtimes of all solvers, except the new one, on all instances.
There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
This allows for both comparing the solvers as well as identifying solvers with similar behavior if looking at individual predictions per instance.
%JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking? "similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
%TF: made it clearer

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
%TF: I still find it worth mentioning. So, I added it back to the table.
Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{comparability}).
%JB: we should rename the "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
%TF: Renamed it to "comparability"
Finally, our strategy incrementally selects only a single benchmark set to compare one new solver $\hat{a}$ with the set of existing solvers $\mathcal{A}$ whose cardinality can be any number (\emph{scalability}).
%JB: for what do we measure cost here?
%TF: made it clearer
We avoid comparing pairs of solvers, which requires a non-constant number of benchmark sets.
%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if the binary-search-/recursive-sorting-like procedure used
%TF: "non-constant number of benchmark sets"
Since a new solver idea is rarely best on the first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

To sum up our contribution, we take the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022} (5355 instances, full runtimes of 28 solvers), perform cross-validation by treating each solver as new once, and predict these solvers' PAR-2 rank with about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
Both results are averages across all cross-validation iterations.
%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)
%TF: added a bit more information

For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
%JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
%TF: it's supposed to be a short overview of what is to come
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
%JB: as commented above, we should either remove this criterion or show it in the table
%TF: added it back to the table

@@

Benchmark-set selection is an important issue for solver competitions, e.g., the SAT Competition.
In such competitions, the organizers define the rules for composing the corresponding benchmarks.
Balint et al.~\cite{balint2015overview} provide an overview of benchmark-selection criteria in different solver competitions.
%JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al. (2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
%TF: @Markus maybe you are better at briefly discussing the  SAT Competition 2022 and previous competitions
Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
M{\i}s{\i}r~\cite{misir2021benchmark} presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances.
The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisfied criteria
%TF: I still find it useful though to discuss everything in the table. Let's cut space in the end.
Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, solvers are \emph{comparable} within their approach in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.

@@

\begin{table}[tbp]
  \centering
  \caption{
  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
  }
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Comparability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

\subsubsection{SAT-solver configuration.}
Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
The goal is to tune SAT solvers for a given sub-domain of problem instances.

SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
%JB: discussion of SMAC is too long, we should cut it a bit
%TF: Agreed. Not sure however where. I find it useful to discuss the similarities and differences between algorithm configuration and our problem.
Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
There are mainly two similarities.
First, both approaches are model-based.
While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already-known runtimes of instances and runtime information on already-known solvers.
Second, both approaches employ some kind of selection strategy that is based on the previously mentioned model.
Theirs selects configurations rather than instances though.
Furthermore, Hutter et. al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with few data points) to select the next configuration for which to acquire runtime information.
SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.

While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to the best-performing solvers (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
%TF: made it clearer
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
And third, configurations within their approach are not \emph{comparable} since algorithm configuration seeks to find solemnly the best performing configuration.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)
%TF: changed

\subsubsection{Incremental benchmark-set-selection.}
Matricon et.~al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
The main difference is that Matricon et.~al. compare a solver to another one, while we compare it to a set of existing solvers.
To the best of our knowledge, we are the first to tackle the latter problem.
%JB: again, I think there are too many details for describing this single approach
%TF: unsure what to cut as the concrete approaches are also used later in the main section

Given a pair of SAT solvers, Matricon et.~al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
%Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et. al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
Their experiments show that random instance selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
%JB: in our evaluation, we should state whether our results are consistent with this or not
%TF: will do

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes solvers partially \emph{comparable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
As they compare only pairs of solvers, evaluating a new solver requires a non-constant number of benchmark sets, which in turn requires more runtime.
%JB: maybe rather log-linear than quadratic, see a prior comment in intro
%TF: changed

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: not sure if this dilemma arises here as well; a dilemma is typical for searching for one optimal solution in continuous space, where there is an incentive to exploit neighborhoods in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
%TF: removed the exploration-exploitation dilemma

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.25\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {\large Train. Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {\large $x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {\large Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {\large Approx. Func.};
  \node at (-2.8,0.25) {\large $f$};
  \end{tikzpicture}
  }
  \caption{Passive}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {\large Train. Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {\large $x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {\large Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {\large Approx. Func.};
  \node at (-2.8,0.25) {\large $f$};
  \node[right] at (-1.2,2.4) {\large Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {\large $x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.35\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {\large Train. Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {\large $x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {\large Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {\large Approx. Func.};
  \node at (-2.8,0.25) {\large $f$};
  \node[right, label={[align=left]\large Gen.\\\large Model}] at (-0.6,1.85) {};
  %JB: does the model need to be a GAN or can it also be a different type of model?
  %TF: not necessarily, changed it
  \node[right] at (-1.3,3.135) {\large $x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}

  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15})}
  %JB: slightly larger font size in the figure (at least having the size of the caption) would be nice
  %TF: made it larger
  \label{fig:learning}
\end{figure}
